Identifying regime transitions for water governance at a basin scale

### 2022-04-29 paper/main/01_introduction.tex
docs(finish-appendix): full version, with finished appendix, well structured main text, all references, correct grammar
BREAKING CHANGE:

Water governance refers to the political, social, economic, and administrative systems that influence the use and management of water \cite{oecd2018, wang2017}.
For populated large river basins, missing governance means missing sustainability, and a first important step in understanding the transitions with successful water governance is identifying the different regimes \cite{kjellen2015, grafton2013}.
Regimes of water governance maintained by concreted intertwine within human-water systems (such as management, institutions, and exploitations), as a stable state in structures and functions
\cite{falkenmark2021,bressers2013,loch2020}.
Therefore, regime shifts sometimes lead to new water governance challenges as both signals and consequences of substantive changes in human-water systems.
The lack of a comprehensive but straightforward approach to identifying changes in water governance regimes challenges sustainability, and filling this gap can well align human and water systems (Figure~\ref{fig:framework}).

>>>>

Water governance refers to the political, social, economic, and administrative systems that influence the use and management of water \cite{oecd2018, wang2017}.
For populated large river basins, missing governance means missing sustainability, and a first critical step in understanding the transitions with successful water governance is identifying the different regimes \cite{kjellen2015, grafton2013}.
Regimes of water governance maintained by concreted intertwine within human-water systems (such as management, institutions, and exploitations) as a stable state in structures and functions
\cite{falkenmark2021,bressers2013,loch2020}.
Therefore, regime shifts sometimes lead to new water governance challenges as both signals and consequences of substantive changes in human-water systems.
The lack of a comprehensive but straightforward approach to identifying changes in water governance regimes challenges sustainability, and filling this gap can well align human and water systems (Figure~\ref{fig:framework}).

### 2022-04-29 paper/main/02_methods.tex
docs(finish-appendix): full version, with finished appendix, well structured main text, all references, correct grammar
BREAKING CHANGE:

	Then, the Integrated Water Governance Index (IWGI) is the average of the normalized indicators $I'_x$:
	\begin{equation}
		IWGI = I'_S + I'_P - I'_A
	\end{equation}

@@

	\subsubsection*{Indicator of stress}
	We used the scarcity-flexibility-variability (SFV) water stress index proposed in Qin et al., 2019 to evaluate water stress \cite{qin2019}. This metric takes into account management measures (such as the construction of reservoirs) and the impact of changes in the structure of water use on the evaluation of water scarcity. Based on the hydrological and economic context of YRB, four main regions are divided (Source Region, Upper Region, Middle Region, and Lower Region, see \textit{\ref{secA1}}). For the whole YRB, the indicator of water stress $I_S$ is the average of all regions' SFV-index:
	\begin{equation}
		I_S = \frac{1}{4} * \sum_{i=1}^4 SFV_{i}
	\end{equation}

	Where $SFV_i$ is the SFV-index for region $i$, and the detailed calculation of $SFV_i$ can be found in the \ref{secA2}.

@@

	\subsubsection*{Indicator of allocations}
	%$AEM$ 指标度量的是水资源配置“不均匀”的程度，类似于信息熵是对“混乱程度”的度量，即参与分配的各个单位之间，分配比例差距越大，则熵越小。
	% 而我们的指标应反映随着社会发展，水资源配置在区域之间更加均衡（比例差距小）、整体满足不同用水部门的发展需求（比例差距减小）、但不同区域存在部门分工（比例差距增大）的趋势。
	To describe allocations $I_A$, we designed an indicator based on entropy, called Allocation Entropy Metric (AEM), which measures the degree of evenness in water allocation:

@@

	where $p_{i}$ is the water proportion of region $i$ to the whole basin ($N=4$ regions in sum). Based on the hydrological and economic context of YRB, four main regions are divided (Source Region, Upper Region, Middle Region, and Lower Region, see \textit{\ref{secA1}}).

	\subsection{Change points detection}
		With no assumptions about the distribution of the data, we applied the Pettitt (1979) approach of change-point detection to detect a single change-point in hydrological time series with continuous data
        \cite{pettitt1979}.
		It tests $H0$: The variables follow one or more distributions with the exact location parameter (no change) against the alternative: a change point exists.
		Mathematically, when a sequence of random variables is divided into two segments represented by $\mathrm{x}_{1}, \mathrm{x}_{1}, \ldots, x_{t_{0}} \text{and} x_{t_{0}+1}, x_{t_{0}+2}, \ldots, x_{T}$, if each segment has a common distribution function, i.e., $F_1(x)$, $F_2(x)$ and $F_1(x) \neq F_2(x)$, then the change point is identified at $t_0$. To achieve the identification of change point, a statistical index $U_{t,T}$ is defined as follows:

@@

		%实际上这个阈值取0.001至0.1不会影响我们的结果，我们识别的断点是相当稳定的
		We used $\alpha = 0.001$ as the threshold level of the p-value, meaning that the probability of a statistically significant change-point judgment being valid was more than $99.9\%$. We divided the series into two at that point and analyzed each series separately until all significant change points were detected. Though two breakpoints were detected in the main text with $\alpha = 0.001$, in fact the threshold from $0.0005$ to $0.05$ does not affect our results, and the breakpoints we identified are fairly robust (see \textit{SI Appendix} Figure~\ref{fig:sensitivity}).

	% % 计算贡献度
	% \subsection{Contribution decomposition}
	% 	We decomposed the amount of variation in each index at different periods to quantify each influence's contribution to the index. Using the Integrated Water Governance Index (IWGI) as an example, its value is influenced by normalized indicators in three dimensions: stress ($I'_S$), purpose ($I'_P$) and allocation ($I'_A$). We can calculate their differences between two certain years ($y_2$ and $y_1$, $y_2 > y_1$) by:
	% 	\begin{align}
	% 	\Delta IWGI &= (I'_{P_{y_2}} + I'_{A_{y_2}} - I'_{S_{y_2}}) - (I'_{P_{y_1}} + I'_{A_{y_1}} - I'_{S_{y_1}}) \\
	% 	&= (I'_{P_{y_2}} - I'_{P_{y_1}}) + (I'_{A_{y_2}} - I'_{C_{y_1}}) + (I'_{S_{y_1}} - I'_{S_{y_2}}) \\
	% 	&= \Delta I'_P + \Delta I'_A + (-\Delta I'_S)
	% 	\end{align}
	% 	Then, the contribution of dimension $x$ to IWGI's changes can be referred to as:
	% 	\begin{equation}
	% 		Contribution_x = \frac{\Delta I'_x}{\lvert \Delta IWGI \rvert}
	% 	\end{equation}

		% Since $Contribution_x$ can be positive or negative, we used an absolute value to evaluate the contribution (in proportion) of a certain dimension in the all three dimensions to the changes:
		% \begin{equation}
		% 	Contribution_x\% = \frac{|\Delta I'_x|}{\sum_x \lvert \Delta I'_x \rvert}
		% \end{equation}

		% or just for contributions (in proportion) within a certain year $j$ to the regime:
		% \begin{equation}
		% 	Contribution_{x,j}\% = \frac{\lvert I'_{x, j} \rvert}{\lvert \sum_x I'_{x, j} \rvert}
		% \end{equation}

	\subsection{Datasets}
	% 为了计算 IWGI ，我们需要计算多个指标及子指标，所有使用的数据集都在表中列出，数据的详细介绍可见补充材料。
	In order to calculate IWGI from 1965 to 2013, we need to calculate multiple indicators and sub-indicators. All the datasets used are listed in the \textit{SI Appendix} table S1. A detailed description of the data is provided in the supplementary materials \textit{SI Appendix} Methods S2.

>>>>

	Then, the Integrated Water Governance Index (IWGI) is an average of the normalized indicators $I'_x$:
	\begin{equation}
		IWGI = (I'_S + I'_P + I'_A) / 3
	\end{equation}

@@

	\subsubsection*{Indicator of stress}
	We used the scarcity-flexibility-variability (SFV) water stress index proposed in Qin et al., 2019 to evaluate water stress \cite{qin2019}. This metric considers management measures (such as the construction of reservoirs) and the impact of changes in the structure of water use on the evaluation of water scarcity. Based on the hydrological and economic context of YRB, four second-level regions are divided (Source Region, Upper Region, Middle Region, and Lower Region, see \textit{Appendix}\ref{secA1}). For the whole YRB, the indicator of water stress $I_S$ is the average of all regions' SFV-index:
	\begin{equation}
		I_S = \frac{1}{4} * \sum_{i=1}^4 SFV_{i}
	\end{equation}

	Where $SFV_i$ is the SFV-index for region $i$, and the detailed calculation of $SFV_i$ can be found in the \textit{Appendix}\ref{secA2}.

@@

	\subsubsection*{Indicator of allocations}
	To describe allocations $I_A$, we designed an indicator based on entropy, called Allocation Entropy Metric (AEM), which measures the degree of evenness in water allocation:

@@

	where $p_{i}$ is the water proportion of region $i$ to the whole basin (here, $N=4$ considering divided regions in the YRB, see \textit{Appendix}\ref{secA1}).

	\subsection{Change points detection}
		With no assumptions about the distribution of the data, we applied the Pettitt (1979) approach of change-point detection to detect a single change-point in hydrological time series with continuous data
        \cite{pettitt1979}.
		It tests $H0$: The variables follow one or more distributions with the exact location parameter (no change) against the alternative: a change point exists.
		Mathematically, when a sequence of random variables is divided into two segments represented by $\mathrm{x}_{1}, \mathrm{x}_{1}, \ldots, x_{t_{0}}$ and $x_{t_{0}+1}, x_{t_{0}+2}, \ldots, x_{T}$, if each segment has a common distribution function, i.e., $F_1(x)$, $F_2(x)$ and $F_1(x) \neq F_2(x)$, then the change point is identified at $t_0$. To achieve the identification of change point, a statistical index $U_{t,T}$ is defined as follows:

@@

		%实际上这个阈值取0.001至0.1不会影响我们的结果，我们识别的断点是相当稳定的
		We used $\alpha = 0.001$ as the threshold level of the p-value, meaning that the probability of a statistically significant change-point judgment being valid was more than $99.9\%$. We divided the series into two at that point and analyzed each series separately until all significant change points were detected. Though two break points in the main text with $\alpha = 0.001$, the threshold from $0.0005$ to $0.05$ does not affect our results, and the breakpoints we identified are robust (see \textit{Appendix} Figure~\ref{fig:sensitivity}).

	\subsection{Datasets}
	% 为了计算 IWGI ，我们需要计算多个指标及子指标，所有使用的数据集都在表中列出，数据的详细介绍可见补充材料。
	In order to calculate IWGI in the YRB, all the datasets we need are listed in the \textit{SI Appendix} Table~\ref{tab:datasets} with a detailed description in \ref{secA3}.

### 2022-04-29 paper/main/03_results.tex
docs(finish-appendix): full version, with finished appendix, well structured main text, all references, correct grammar
BREAKING CHANGE:

\begin{figure*}[th!]
	\centering
	\includegraphics[width=0.9\linewidth]{main/causes.pdf}
	\caption{
		Causes of water governance regime shifts in the YRB: environmental change, economic growth and efficiency changes, social transformation, and governance policies.
		\textbf{A.} Changes in the total irrigated area (orange line) and water use intensity ($WU/A$, water use divided by the area, the green dot line).
		\textbf{B.} Changes in gross values added (GVA) of industry and services (blue line) and their water use intensities ($WU/GVA$ WU divided by the GVA, the red dot line).
		\textbf{C.} Completed time of each new reservoir and their located region's water use (LWU) percentages as a proportion of the total basinal water use (BWU) at that time. Red circles denote the reservoirs mainly for managing and regulating the whole basin.
		The size of each circle indicates the magnitude of its water storage capacity. Some crucial reservoirs include (1) Xiaolangdi reservoir and Sanmen Reservoir, which were constructed mainly for managing sediments; and (2) Impoundments at Liujia Gorge, Longyang Gorge, which was constructed mainly for managing floodwater discharge and water regulation. These named reservoirs are significant for the whole basin rather than local development.
		\textbf{D.} Social transformations (red triangles) and national-level governance policies (the circles, different colours denote signed by different state institutions, see \textit{SI Appendix} Table S2). The light grey bars year-by-year show the frequency of official documents related to the YRB at a basinal scale (see \textit{SI Appendix} Table S3).
	}
	\label{fig:Causes}
\end{figure*}

% 进一步挖掘IWGI变化的根本原因，
The underlying causes of changes in the IWGI are associated with various factors but are different in the two regime shifts.
% 灌溉区扩张和工业和服务业的经济增长是P1和P2目的变化的关键。
Changing water demands and supply were critical to the shift between P1 and P2.
% P1年，黄河流域灌溉农业面积以图A的速度快速扩张，灌溉用水占主导(1965年占总用水的$81.56\%$, 1978年占总用水的$83.17\%$，图S3)。
As the dominant water demand during the massive supply regime (P1), the area of irrigated agriculture in the YRB expanded rapidly at a rate of $0.25*10^6 ha/yr$ (Figure\ref{fig:Causes} A), simultaneously supported by increasing supply through the construction of reservoirs (\textit{SI Appendix} Figure~\ref{fig:reservoirs}).
% 然而，进入P2后，灌溉区扩张停滞，工业和服务业逐渐增长，用水需求增加(图re\ref{fig: Causes} B)，导致灌溉用水量比例下降$8\%$ S3。
Entering the transformation governance regime (P2), however, the expansion of irrigated areas slowed down, and industry and services gradually took off (Figure\ref{fig:Causes} A and B).
% 水分利用效率由P2变化到P3。
Then, the efficiency of water use changed obviously from P2 to P3.
Not only irrigated areas keep their slow expansion in the adaptation oriented regime (P3) (Figure\ref{fig:Causes}A), but industry and urban services also assumed a more vital economic role (represented by Gross Added Values, GVA) (Figure~\ref{fig:Causes}B).
However, because of more efficient technology and better water conservation practices (\textit{SI Appendix} Figure~\ref{fig:wci}), both experienced significant declines in water use for a unit irrigated area or unit production (Figure~\ref{fig:Causes}A and Figure~\ref{fig:Causes}B).
As a result, the differences between sectors and regions in water use reduced while the total water stress steadily remained high during the adaptation oriented regime (Figure~\ref{fig:IWGI}A).

% 最后，环境背景、社会转型和水治理政策在这三个时期都发挥了作用。
Environmental context, social transformation and policies played roles in all three regimes.
We calculated the ratios of regional and basinal water use for each reservoir (R/B ratio) (Figure~\ref{fig:Causes}C), with a higher ratio representing a potential role in water supply rather than basinal regulations.
Under the guiding ethos of ``conquering nature'', most of the reservoirs were built in regions with high water demands during the massive supply regime (R/B ratios were significantly higher, $p<0.01$, see Figure~\ref{fig:Causes}C), when natural water resources were also relatively abundant (\textit{SI Appendix} Figure~\ref{fig:water resources}).
Since the transformation governance regime (P2), the number of new reservoirs decreased significantly and significantly increased basinal policies rigorously controlled the allocation of water (Figure~\ref{fig:Causes}D, $p<0.01$ and \textit{SI Appendix} Figure~\ref{fig:reservoirs}).
In the adaptation oriented regime, authorities proposed more national-level water governance policies under the guidance of the national strategy ``environmental regulation'' (Figure~\ref{fig:Causes}D).
Taken together, the regime shift from P1 to P2 is in line with the increasing water supply and demands; while driven by regulatory policies and efficiency enhancement under stable water stress from P2 to P3.

>>>>

\begin{figure*}[th!]
	\centering
	\includegraphics[width=0.9\linewidth]{main/causes.pdf}
	\caption{
		Causes of water governance regime shifts in the YRB.
		% environmental change, economic growth and efficiency changes, social transformation, and governance policies.
		\textbf{A.} Changes in the total irrigated area (orange line) and water use intensity ($WU/A$, water use divided by the irrigated area, the green dot line).
		\textbf{B.} Changes in gross values added (GVA) of industry and services (blue line) and their water use intensities ($WU/GVA$ WU divided by the GVA, the red dot line).
		\textbf{C.} Completed time of each new reservoir and their located region's water use (LWU) percentages as a proportion of the total basinal water use (BWU) at that time. Red circles denote the reservoirs mainly for managing and regulating the whole basin.
		The size of each circle indicates the magnitude of its water storage capacity.
		% Some crucial reservoirs include (1) Xiaolangdi reservoir and Sanmen Reservoir, which were constructed mainly for managing sediments; and (2) Impoundments at Liujia Gorge, Longyang Gorge, which were constructed mainly for managing floodwater discharge and water regulation. These named reservoirs are significant for the whole basin rather than local development.
		\textbf{D.} Social transformations (red triangles) and national-level governance policies (the circles, different colours denote signed by different state institutions, see \textit{Appendix} Table~\ref{tab:policies}). The light grey bars count official documents related to the YRB on a basinal scale (the Yellow River Events).
	}
	\label{fig:Causes}
\end{figure*}

% 进一步挖掘IWGI变化的根本原因，
The underlying causes of changes in the IWGI are associated with various factors but are different in the two regime shifts.
% 灌溉区扩张和工业和服务业的经济增长是P1和P2目的变化的关键。
Changing water demands and supply were critical to the shift between P1 and P2.
% P1年，黄河流域灌溉农业面积以图A的速度快速扩张，灌溉用水占主导(1965年占总用水的$81.56\%$, 1978年占总用水的$83.17\%$，图S3)。
As the dominant water demand during the massive supply regime (P1), the area of irrigated agriculture in the YRB expanded rapidly at a rate of $0.25*10^6 ha/yr$ (Figure\ref{fig:Causes} A), simultaneously supported by increasing supply through the construction of reservoirs (\textit{Appendix} Figure~\ref{fig:reservoirs}).
% 然而，进入P2后，灌溉区扩张停滞，工业和服务业逐渐增长，用水需求增加(图re\ref{fig: Causes} B)，导致灌溉用水量比例下降$8\%$ S3。
Entering the transformation governance regime (P2), however, the expansion of irrigated areas slowed down, and industry and services gradually took off (Figure\ref{fig:Causes} A and B).
% 水分利用效率由P2变化到P3。
Then, the efficiency of water use changed obviously from P2 to P3.
Not only irrigated areas keep their slow expansion in the adaptation oriented regime (P3) (Figure\ref{fig:Causes}A), but industry and urban services also assumed a more vital economic role (represented by Gross Added Values, GVA) (Figure~\ref{fig:Causes}B).
Because of increased efficiency, however, they both experienced significant declines in water use for a unit irrigated area or unit production (Figure~\ref{fig:Causes}A and Figure~\ref{fig:Causes}B).
As a result, the differences between sectors and regions in water use reduced while the total water stress steadily remained high during the adaptation oriented regime (Figure~\ref{fig:IWGI}A).

% 最后，环境背景、社会转型和水治理政策在这三个时期都发挥了作用。
Environmental context, social transformation and policies played roles in all three regimes.
We calculated the ratios of regional and basinal water use for each reservoir (R/B ratio) (Figure~\ref{fig:Causes}C), with a higher ratio representing a potential role in water supply rather than basinal regulations.
Under the guiding ethos of ``conquering nature'', most of the reservoirs were built in regions with high water demands during the massive supply regime (R/B ratios were significantly higher, $p<0.01$, see Figure~\ref{fig:Causes}C)).
Since the transformation governance regime (P2), the number of new reservoirs decreased significantly and significantly increased basinal policies rigorously controlled the allocation of water (Figure~\ref{fig:Causes}D, $p<0.01$ and \textit{SI Appendix} Figure~\ref{fig:reservoirs}).
In the adaptation oriented regime, authorities proposed more national-level water governance policies under the guidance of the national strategy ``environmental regulation'' (Figure~\ref{fig:Causes}D).
The regime shift from P1 to P2 is in line with the increasing water supply and demands; while driven by regulatory policies and efficiency enhancement under stable water stress from P2 to P3.

### 2022-04-29 paper/main/04_discussion.tex
docs(finish-appendix): full version, with finished appendix, well structured main text, all references, correct grammar
BREAKING CHANGE:

\begin{figure*}[htbp!]
	\centering
	\includegraphics[width=0.9\linewidth]{main/transition.png}
	\caption{
		Transition schema in hydrosocial cycle and water governance regimes. The natural water cycle dominates blue pathways, while socio-economic feedback dominates red pathways.
		\textbf{A.} As socio-economic systems develop, non-provisioning water demand increases; simultaneously, increased adaptive capacity by engineering allows people to manage water resources to alleviate the water stress.
		\textbf{B.} With further human interventions, trade-offs between provisioning-purpose and non-provisioning water use become prominent; a basin-wide socio-economic system requires more organized water governance.
		Thus, \textbf{C. the hydrosocial water cycle transition} correlates with the water governance regime shifts (in the YRB, they are massive supply regime, transformation governance regime, and adaptation oriented regime). The transformation governance regime shift occurs following the water deficit, with the rapid growth of adaptive capacity.
		\textbf{D. Water governance challenges} Through the transitional regimes, water governance faces primarily economic and environmental challenges but social and policy challenges later.
	}
	\label{fig:summary}
\end{figure*}

@@

% 总体而言，长江流域的水治理是“改善供给、转变治理、增强适应”的水社会循环总体转型中最突出的例子之一。
In general, water governance of the YRB is among the most prominent example in the general transition of hydrosocial cycle -``improving supply, transforming governance, and enhancing adaptation''.
With each dimension changing gradually, the emergence of regime shifts drives the water governance challenges at a basin-scale: primarily economic and environmental before the transformation but social and policy-related towards the end (Figure~\ref{fig:summary}) \cite{singh2019,porcher2019}.
In the analogy at a global scale, the resource challenges, represented as water shortage and water supplying difficulties, are mainly faced by undeveloped and developing basins
\cite{allan2019,speed2013,liu2012a}.
Alternatively, highly-controlled and developed basins (especially for transboundary rivers) must mainly resolve structural challenges, such as water disputes or lack of equity, and may be in urgent need of novel flexible, efficient sociopolitical governance structures
\cite{unep-dhi2016,mirumachi2015}.
Linking regime shifts to the governance challenges, the implementation of IWGI thus offers a comprehensive and straightforward way to interpret the intertwines between water governance and the hydrosocial transition.

%* limitations & direction
One of the main limitations in the approach is the few data worldwide with a long-term period, which means still a gap between the comprehensively identifying and widespread application of IWGI.
However, we assumed that all water governance issues are relative to ``who gets water, when and how'', so water stress, purposes of water services, and water allocation patterns matter.
Therefore, we suggest that choices of the indicators for the dimensions can be adapted according to available datasets as the intertwines between the underlying components are much more crucial in understanding the transition of governance regimes holistically.
In today's world, the regime shifts from biophysical to hydrosocial control of dynamics may become increasingly widespread, so the comprehensive strategies to address governance challenges have become the core of complex human-water systems
\cite{cumming2018,cumming2014,jaeger2019}.
Although river basins have shown improvements in water management technologies and water use efficiency, many of them are still approaching planetary boundaries where human-water systems may collapse
\cite{gleeson2020, wang-erlandsson2022}.
A deeper understanding of governance that incorporates ideas of non-linear regime shifts and transformations should help shift the focus of governance towards maintaining the resilience of the basin’s social-ecological system and improving its sustainability
\cite{falkenmark2019}.

>>>>

\begin{figure*}[htbp!]
	\centering
	\includegraphics[width=0.8\linewidth]{main/transition.png}
	\caption{
		Transition schema in hydrosocial cycle and water governance regimes. The natural water cycle dominates blue pathways, while socio-economic feedback dominates red pathways.
		\textbf{A.} As socio-economic systems develop, non-provisioning water demand increases; simultaneously, increased adaptive capacity by engineering allows people to manage water resources to alleviate the water stress.
		\textbf{B.} With further human interventions, trade-offs between provisioning-purpose and non-provisioning water use become prominent; a basin-wide socio-economic system requires more organized water governance.
		Thus, \textbf{C. the hydrosocial water cycle transition} correlates with the water governance regime shifts (in the YRB, they are massive supply regime, transformation governance regime, and adaptation oriented regime). The transformation governance regime shift occurs following the water deficit, with the rapid growth of adaptive capacity.
		\textbf{D. Water governance challenges} Through the transitional regimes, water governance faces primarily economic and environmental challenges but social and policy challenges later.
	}
	\label{fig:summary}
\end{figure*}

@@

% 总体而言，长江流域的水治理是“改善供给、转变治理、增强适应”的水社会循环总体转型中最突出的例子之一。
In general, water governance of the YRB is among the most prominent example in the general transition of hydrosocial cycle -``improving supply, transforming governance, and enhancing adaptation''.
With each dimension changing gradually, the emergence of regime shifts drives the water governance challenges at a basin-scale: primarily economic and environmental before the transformation but social and policy-related towards the end (Figure~\ref{fig:summary}) \cite{singh2019,porcher2019}.
In the analogy at a global scale, the resource challenges, represented as water shortage and water supplying difficulties, are mainly faced by undeveloped and developing basins
\cite{allan2019,speed2013,liu2012a}.
Alternatively, highly-controlled and developed basins (especially for transboundary rivers) must mainly resolve structural challenges, such as water disputes or lack of equity, and maybe in urgent need of novel flexible, efficient sociopolitical governance structures
\cite{unep-dhi2016,mirumachi2015}.
Linking regime shifts to the governance challenges, the implementation of IWGI thus offers a comprehensive and straightforward way to interpret the intertwines between water governance and the hydrosocial transition.

%* limitations & direction
One of the main limitations in the approach is the few data worldwide with a long-term period, which means still a gap between the comprehensively identifying and widespread application of IWGI.
However, we assumed that all water governance issues are relative to ``who gets water, when and how'', so water stress, purposes of water services, and water allocation patterns matter.
Therefore, we suggest that choices of the indicators for the dimensions can be adapted according to available datasets as the intertwines between the underlying components are much more crucial in holistically understanding the transition of governance regimes.
In today's world, the regime shifts from biophysical to hydrosocial control of dynamics may become increasingly widespread, so the comprehensive strategies to address governance challenges have become the core of complex human-water systems
\cite{cumming2018,cumming2014,jaeger2019}.
Although river basins have shown improvements in water management technologies and water use efficiency, many of them are still approaching planetary boundaries where human-water systems may collapse
\cite{gleeson2020, wang-erlandsson2022}.
A deeper understanding of governance that incorporates ideas of non-linear regime shifts and transformations should help shift the focus of governance towards maintaining the resilience of the basin’s social-ecological system and improving its sustainability
\cite{falkenmark2019}.
